---
title: "Good Coding Practices"
author: "Andrew Do"
date: "July 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
sol = TRUE
```

```{r}
library(dplyr)
```

## Short Exercises
Rewrite the following according to our class style guide:

1.
```{r, eval = FALSE}
x < - c( NA,1,-2,3,-4,5 )
  y <- 3
  y <- x * - 1
  y [ y>0 ]
```

```{r, echo = sol}
x <- c( NA,1,-2,3,-4,5 )
  y <- 3
  y <- x * - 1
  y [ y>0 ]
```

2. 
```{r}
mean <- function(x) {
sum (x)/length( x )
}
```

```{r, echo = sol}
Mean <- function(x){
sum (x)/length( x )
}
```

3. 
```{r}
library( ggplot2 )
second_version_of_mpg <- mpg[ mpg$cyl == 6,]
second_version_of_mpg$class <- as.character(second_version_of_mpg$class)
ggplot(data=second_version_of_mpg,aes(cty,hwy,color=class))+geom_point()+geom_smooth( method = "lm", se=FALSE ) + labs (title="City vs. Highway MPG" )
```

```{r, echo = sol}
mpg_6_cylinders <- mpg[ mpg$cyl == 6,]
mpg_6_cylinders$class <- as.character(second_version_of_mpg$class)
ggplot(data=mpg_6_cylinders,aes(cty,hwy,color=class))+
  geom_point()+
  geom_smooth( method = "lm",
               se=FALSE ) + 
  labs (title="City vs. Highway MPG" )
```

## Longer Exercises

Fix the following code with the following goals:

* Readability - an untrained person should be able to guess what's going on
    + Use meaningful names
    + Use whitespace to make the code flow
    + Provide documentation
    + Use functions from the packages we've covered (they're designed to be readable!)
  
* Compactness - do not be more verbose than you have to be
    + Think of ways to vectorize the function
    + Avoid copy-paste
    + Use the `lapply` family
    + Modularize your code (small functions instead of massive ones)
  
* Usable as intended
    + If necessary, include error messages
    + Don't force the user to rely on implicit coercion
    + Output the intentional data type

### Exercise 1
Infer what the function `d` is doing and rewrite it according to our class style guide.  Include documentation so that future users won't have to decipher the code.
```{r}

d <- function(f, x, y, s) {
  stuff<-seq(x,y,by=s)
  stuff2<-c()
  for(i in 1:length(stuff)){thing<-f(stuff[i]); stuff2[i]<-thing}
  m<-max(stuff2)
  mm<-stuff[stuff2==m]
  mm<-min(mm)
  return(list(m,mm))
}

# Example usage
g <- function(x) -(x ^ 2)
d(g, -1, 1, .1)
```

```{r, echo = sol}
MapFunction <- function(f, x, y, s) {#Is this a Map function?
  sequence_from_x_to_y<-seq(x,y,by=s)#sequence
  stored_vector_for_map<-c()#empty vector
  for(i in 1:length(sequence_from_x_to_y)){#for loop
     mapped_item<- f(sequence_from_x_to_y[i])
     stored_vector_for_map[i]<- mapped_item
     }#Maps the function from stuff, and puts the mapped output into the empty vector
  max_mapped_vector<- max(stored_vector_for_map)#returns the max of the stored_mapped_items
  index_vector<- sequence_from_x_to_y[stored_vector_for_map==max_mapped_vector]#returns the index from the original sequence of the map item
min_index<- min(index_vector)#
  return(list(max_mapped_vector,index_vector))#returns a list of the above functions
}

# Example usage
g <- function(x) -(x ^ 2)
MapFunction(g, -1, 1, .1)
```


### Exercise 2
Figure out what the `ss` function is doing, rewrite it so it uses more expressive functions according to our style guide.  Provide documentation.
```{r}
ss <- function(x) 
{x<-x[!is.na(x)]#returns the input x with all of the NA values removed.
a<-x[1];for(i in 2:length(x)) a[i]<-(a[i-1]+x[i])#the first item of the new vector x
a<-a/(1:length(x))#divides every element in the list by ???
foo<-x#creates a mirror of x in another variable
cc<-0
for(i in 1:length(foo)){x<-foo[1:i]
c<-sort(x);
if(length(x)%%2==0){cc[i]<-(c[floor((length(c)+1)/2)] + c[ceiling((length(c)+1)/2)])/2} else{ cc[i]<-c[(length(c)+1)/2]}}
return(list(a,cc))
}

# Example usage
ss(c(5,10,15,20,5,10,15,20))
ss(1:10)
ss(c(1,2,6,2,4,5))
```

```{r, echo = sol}
ss <- function(vector) 
{vector<-vector[!is.na(vector)]#returns the input x with all of the NA values removed.
first_item_vector<-vector[1];#the first item of the new vector x
for(i in 2:length(vector))
  { first_item_vector[i]<-(first_item_vector[i-1]+vector[i])
  }#for every 2nd item after.
first_item_vector<-first_item_vector/(1:length(vector))#divides every element in the list by ???
vector_copy<-vector #creates a mirror of x in another variable
cc<-0 #This is a bagel of the highest quality
for(i in 1:length(vector_copy)){vector <- vector_copy [1:i]}#
sorted_vector<-sort(vector)#the sorted vector of x;
if(length(vector)%%2== 0)
  {cc[i]<- (sorted_vector[floor((length(sorted_vector) + 1) / 2)] + 
             sorted_vector[ceiling((length(sorted_vector)+1)/2)])/2}#takes the median of vector c if length is even
  else{ cc[i]<-vector[(length(vector)+1)/2]}#takes the median of vector c if length is odd
return(list(first_item_vector,cc))
}

# Example usage
ss(c(5,10,15,20,5,10,15,20))
ss(1:10)
ss(c(1,2,6,2,4,5))
```

### Exercise 3
Rewrite the following script in a more compact form using packages we've covered in this class.  It shouldn't take more than 7 function calls---3 if you took the time to read the `dplyr` vignette!
```{r}
d<-iris
dd<-unique(d[,5])
d1<-d[which(d[,5]==dd[1]),]
d2<-d[which(d[,5]==dd[2]),]
d3<-d[which(d[,5]==dd[3]),]
d1[,1] <- mean(d1[,1])
d1[,2] <- mean(d1[,2])
d1[,3] <- mean(d1[,3])
d1[,4] <- mean(d1[,4])
d2[,1] <- mean(d2[,1])
d2[,2] <- mean(d2[,2])
d2[,3] <- mean(d2[,3])
d2[,4] <- mean(d2[,4])
d3[,1] <- mean(d3[,1])
d3[,2] <- mean(d3[,2])
d3[,3] <- mean(d3[,3])
d3[,4] <- mean(d3[,4])
ddd<-rbind(d1[1,],d2[1,],d3[1,])
rownames(ddd) <- NULL

```

```{r, echo = sol}
iris_summary_df<-iris %>% 
  group_by(Species) %>% 
  summarise(S_Length=mean(Sepal.Length),S_Width=mean(Sepal.Width),P_Length=mean(Petal.Length),P_Width<-mean(Petal.Width))
iris_summary_df<- iris_summary_df[c(2,3,4,5,1)]
```

## Remote Computing Practice
Upload the `.rmd` file to `Documents/lab19/` on your server account and knit it using the command line. Submit the `.html` file produced by knitting on the server.  Assuming you're in the folder with the markdown file on the server, the command is:

```{sh, eval = FALSE}
Rscript -e "rmarkdown::render('lab19.Rmd')"
```


```{r, echo = sol}
sessionInfo()
```

