---
title: "hw8"
author: "Due: 11:59pm Saturday 8/6/2016.  Instructors stop answering questions at 5pm."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(dplyr)
library(ggplot2)
library(animation)
solutions<-readRDS("hw8_sql_results.rds")
library(nycflights13)
library(purrr)
library(tidyr)
```

### purrr

In this exercise, you'll be using functional programming ideas to simulate the central limit theorem on coin flips.  Using neither loops nor the apply functions, write a simulation pipeline that:

* Generates 1000 samples of each 1, 2, ..., 99, 100 fair coin flips.  So for example, you would simulate flipping 1 coins 1000 times, 2 coins 1000 times, etc.
* For each of the 1, 2, ..., 100 flips, produces a histogram of the proportion of heads.  To be explicit: you will create a histogram for the 1000 proportions calculated from the 1 coin flips.  Then you'll do the same for 2 coin flips, 3, etc.

In your plots
* The x-axis should be limited to 0-1
* The y-axis should be limited to 0-15


```{r, eval = FALSE}
individual_flip = function(x) {
  s = sample(c(1.00,0.00), size = x, replace = TRUE, prob = c(0.5,0.5))
  prop = sum(s)/x
  return(prop)
}

num_of_flips <- 1:100
results <- num_of_flips %>% map(function(x) rerun(1000,individual_flip(x)))


plot_flips <- function(i) {
  coin <- as.data.frame(results[[i]]) %>%
    gather(key = Count,
           value = Proportion) %>% 
    mutate(count = c(1:1000))
  
  ggplot(coin) +
    aes(x = Proportion, y = ..density..) +
    geom_histogram(bins = 30) +
    labs(x = "Proportion of Heads",
         y = "Density",
         title = "Distribution of Heads") +
    scale_x_discrete(limits = c(0,1))+
    scale_y_continuous(limits = c(0,15)) 
    
}

num_of_flips %>% map(plot_flips)
```

Now that you have all the static images, you could "animate" the progression by turning the sequence of images into a gif.  This can be achieved with the `animation` package in R or the standalone `ImageMagick` utility.  This is just an FYI, we will not be looking for a GIF.  Also, Please do NOT upload the images to GitHub nor bCourses.  We'll just be checking your code.

## SQL

These exercises are meant to help you wrap your head around SQL.  Answer the questions below using SQL queries.  You can check your answer using `dplyr` methods, but your grade will be based on the SQL commands.  The `nycflights13` data set has been put into a SQLlite database for these problems.  For documentation about the variable names, see

```{r, eval = FALSE}
help(package = "nycflights13")
```

```{r}
flights <- src_sqlite("nycflights13.mysqlite3")

```


1. Find the full name of each airline that flew to Dallas Fort-Worth (DFW).
```{r}
flights %>% tbl(sql("
  SELECT DISTINCT
    name AS Airline_Name
  FROM 
    nyairlines JOIN nyflights
    ON nyairlines.carrier= nyflights.carrier
  WHERE 
    dest GLOB 'DFW'
                    ")) %>% 
  collect(n=Inf)
```

2. Make a table containing the tail number, year of manufacture, model, number of engines, and number of seats of the planes flown by United Airlines (UA).  Sort the results by year manufactured.

```{r}
flights %>% tbl(sql("
  SELECT 
      DISTINCT nyplanes.tailnum AS tailnum,
      nyplanes.year AS year,
      model,
      engines,
      seats
  FROM nyplanes INNER JOIN nyflights
    ON nyplanes.tailnum = nyflights.tailnum
  WHERE carrier GLOB 'UA'
  ORDER BY year,model
")) %>% collect(n=Inf)
```

3. Make a table with the following information: full airport name in a column called `Airport` and number of flights from New York City to those airports in 2013 in a column called `NumberOfFlights`.  Sort the results in descending order by flight counts.
```{r}
flights %>% tbl(sql("
    SELECT nyairports.name as Airport,
      COUNT(*) as NumberOfFlights
    FROM nyflights JOIN nyairports
    ON nyflights.dest = nyairports.faa 
    GROUP BY Airport
    ORDER BY
      NumberOfFlights DESC
"))%>% 
  collect(n=Inf)
```

4. Find the wind speed during the hour of the sceduled departure time of every flight that had a departure delay of more than 30 minutes.  For example, if a flight was scheduled to leave at 5:59am, you would check the wind speed for 5:00am.  In practice, you would of course be more careful than this.
```{r}
flights %>% tbl(sql("
    SELECT  wind_speed, nyflights.hour ,nyflights.time_hour
    FROM nyflights JOIN nyweather
    ON nyflights.time_hour=nyweather.time_hour 
      AND nyflights.origin=nyweather.origin
    WHERE dep_delay>30
")) %>% 
  collect(n=Inf)
```



