---
title: "lab25"
name: "Kevin Khuu"
date: "July 26, 2016"
output: html_document
---

```{r}
library(dplyr)
```

Write a function called `moveCars()` that moves the system one full step; that is, the function should move the red cars east once and the blue cars north once, with red cars going first.

Your function should take the following inputs:
`m`: a matrix

Your function should return a list containing two elements:
`m`: an updated matrix
`not_gridlocked`: a logical value that should be TRUE if the system changed and FALSE if notx`

```{r}
moveCars<- function(m){
  original<-m
  shiftright_index<- c(ncol(m),1:ncol(m)-1)
  redcars<-m*(m==1)#checks for red cars
  shiftrightredcars<-redcars[,shiftright_index]
  blockedcars<-(m!=0)*1
  m*(m!=1)
  notblockedcars<-(m==0)*1
  
  moveNorth<- function(m){}
  notgridlocked<- original==final
  return(list(final,notgridlocked))
}

```


Check:
```{r,eval=FALSE}
m <- matrix(c(1,0,0,0,2,1,2,0,0,2,0,1,1,0,1,2,0,1,2,1,0,0,1,1,2), nrow = 5, ncol = 5, byrow = T)
identical(moveCars(m), list(matrix(c(0,1,0,0,2,1,2,0,0,2,1,1,0,1,0,2,0,1,2,1,0,0,1,1,2), nrow = 5, ncol = 5, byrow = T), TRUE))

m <-  matrix(c(0,1,0,2,1), nrow = 1, ncol = 5, byrow = T)
identical(moveCars(m), list(matrix(c(1,0,1,2,0), nrow = 1, ncol = 5, byrow = T), TRUE))

m <- matrix(c(2,0,1,2,0), nrow = 5, ncol = 1, byrow = T)
identical(moveCars(m), list(matrix(c(0,0,1,2,2), nrow = 5, ncol = 1, byrow = T), TRUE))
```

Now write a function called `initGrid()` that randomizes the initial system.

Your function should take the following inputs:
`r`: number of rows in the matrix
`c`: number of columns in the matrix
`d`: density of the grid 

Your function should return the following:
`m`: an rxc matrix with entries 0 (no cars), 1 (red cars), or 2 (blue cars)

```{r}
initGrid<-function(r,c,d){
  matrix_contents=sample(c(0,1,2),size=r*c,prob=c(1-d,d/2,d/2),replace=TRUE)
  test_matrix <- matrix(matrix_contents,nrow=r,ncol=c)
  return(test_matrix)
}
```






