---
title: "lab6"
author: "Kevin Khuu"
date: "June 15, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Population Variance

Write a function `MyVar` that calculates the population variance of a vector.  If you need a refresher of what that means, [this section of wikipedia should help](https://en.wikipedia.org/wiki/Variance#Population_variance_and_sample_variance).
```{r}
test_population_variance<-function(x){
  y=mean(x)
  z=(y-x)**2
  return(sum(z)/length(z))
}
```


## Finding Roots
Write a function `zeroes` that finds the roots (zeroes) of a degree two polynomial.  The function should take the coefficients of the polynomial $a^2x + bx + c$ as its arguments and returns a vector of roots.  If there are no real roots, the function should stop and produce an error message: `"This polynomial has no real roots."`
```{r}
zeroes<- function(a,b,c){
  discriminant<- (b**2-4*a*c)
  if (discriminant <0){
    stop("This polynomial has no real roots")
  }else if(discriminant==0){
    return ((-b+discriminant**(1/2))/2*a)
  }else{
    positive_root=(-b+discriminant**(1/2))/2*a
    negative_root=(-b-discriminant**(1/2))/2*a
    return(c(positive_root,negative_root))
  }
}
```


Test your function:
```{r, eval = F}
zeroes(1, 0, 0)
zeroes(1, 0, 1)
zeroes(1, 1, 1)
zeroes(1, 0, -1)
```

## Temperature Conversion

Write a function `ConvertTemperature` that takes a numeric vector of temperatures in Fahrenheit and converts them into Celsius, Kelvin, Rankine, Delisle, Newton, Reamur, or Romer, depending on what the user specifies in an argument called `to`.  If they specify some other temperature scale, return the message "The [their input] temperature scale is not supported".  If you don't know the conversions, (see this table)[https://en.wikipedia.org/wiki/Conversion_of_units_of_temperature#Fahrenheit]
```{r}
ConvertTemperature<-function(x,to){
  if (to =="Celsius") {
  return ((x-32)*(5/9))
} else if (to =="Kelvin") {
  return ((x+459.67)*5/9)
} else if (to =="Rankine"){
  return ((x+459.67))
} else if (to =="Delisle") {
  return((212-x)*(5/6))
} else if (to =="Newton"){
  return ((x-32)*(11/60))
} else if (to =="Reamur") {
  return ((x-32)*(4/9))
} else if (to =="Romer"){
  return ((x-32)*(7/24)+7.5)
} else{stop(str_c("The",to,"temperature scale is not supported",sep=" "))}
}
```

Test your function
```{r, eval = T}
ConvertTemperature(32, "Celsius") == 0
ConvertTemperature(212, "Kelvin") == 373.15
ConvertTemperature(10, "Rankine") == 501.67
```


## Rolling Dice

For this problem, you'll need the `sample` and `pmax` function.  `sample` generates a random sample of a specified size either with or without replacement.  You can specify probability weights for each element of your sample space using the `prob` argument.  By default, each element will have equal probability.
```
sample(x = vector_to_be_sampled, size, replace = FALSE, prob = NULL)
```

```{r}
dice=c(1,2,3,4,5,6)
dice_sample=sample(dice,2,TRUE)
dice_sample
```

1. What does `pmax` do? How is it different from `max`? Hint: test both on the two vectors `c(1, 5, 3)` and `c(2, 4, 2)`. 'pmax' returns 
2. When you "roll dice with advantage", you roll a pair of dice and take the higher value as your roll.  Write a function `AdvRoll` the simulates rolling with advantage `n` times. The function should take on 2 arguments `sides` and `n` and output a length `n` vector of rolls.
```{r}
AdvRoll<-function(sides,n){
  dice_1<-sample(1:sides,n,TRUE)
  dice_2<-sample(1:sides,n,TRUE)
  return(pmax(dice_1,dice_2))
}
```
3. Write your own summary function called `MySummary` that, given a numeric vector `x`, returns the following summary statistics: mean, median, mode, sum, and sum of squares.
```{r}
MySummary<- function(x){
  mode<- function(x){
    return(which.max(table(x)))
  }
  return(c("mean"=mean(x),"median"=median(x),"mode"=mode(x),"sum"=sum(x),"sumsquares"=sum(x**2)))
}
```
4. Using your two functions, simulate "rolling dice with advantage" 1000 times and find the summary statistics.
```{r}
dice_1000_sample=AdvRoll(6,1000)
MySummary(dice_1000_sample)
```


