---
title: "lab5finished"
author: "Andrew Do"
date: "June 13, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(stringr)
library(readr)
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Help for Packages

You can see what an entire package has to offer with
```{r}
help(package = "stringr")
```


## Counting number of vowels

Consider the following vector `letrs` which contains various letters:
```{r}
# Garbage string
letrs <- "gfngenodgsyctyviqyyigeinyngbkwjdjqkrrollhpxdkfldlytympmcgmjynihusebtq"
```

If you were to count the number of vowels in `letrs` you would get the following counts:

- a: 0
- e: 3
- i: 4
- o: 2
- u: 1

This exercise will guide you through how to arrive at this frequency table.

```{r}
# Create a character vector of vowels.  Give it an appropriate name.
vowels<- c("a","e","i","o","u")

# Use str_count to get the counts.  Assign this an appropriate name.
letrs_vowel_count<- str_count(letrs,vowels)
letrs_vowel_count
# Give each element of the vector a name using the names function.
names(letrs_vowel_count)<-c("a","e","i","o","u")
letrs_vowel_count
# The result should look like:
# a e i o u
# 0 3 4 2 1
```

## Exploring new functions: str_pad
Write a short description of what each of the following calls of `str_pad` is doing.  The last three examples were found at the bottom of the function's documentation.

1. In this call of `str_pad`, This appends the input string "a" with character "x" until the combined string and new characters reach width 4.
```{r}
str_pad("a", width = 4, side = "left", pad = "x")
```

2. In this call of `str_pad`, spaces are added to the left of the string until the string reaches width 3.
```{r}
str_pad("a", width = 3, side = "left")
```

3. In this call of `str_pad`, spaces are added to the length.These three strings are of increasing length, so less spaces are added to the third string compared to the first string because the first string was originally shorter than the third string.
```{r}
str_pad(c("a", "abc", "abcdef"), 10)
```

4. In this call of `str_pad`, different instances of the input string are padded with spaces until they reach width 5,10, and 20 respectively.
```{r}
str_pad("a", c(5, 10, 20))
```

5. In this call of `str_pad`, different characters are added to the left of the input string until the each respective instance of the string reaches the width 10.
```{r}
str_pad("a", 10, pad = c("-", "_", " "))
```
## Regular Expressions

We read in the data from the csv file and extract the names column
```{r}
flags <- read_csv("flags.csv")
countries <- flags$name
```

You may have noticed that in the `countries` vector, words in a country's name are separated by `-`.  Use `str_replace_all` to replace the hyphens with spaces.  For example `"Vatican-City"` should be replaced with `"Vatican City"`

```{r}
str_replace_all(countries,"-"," ")
```

Use `str_detect` to see if there are any countries with the world "island" in them, ignoring case.  You may also want to look up how to use `ignore.case`.

```{r}
str_detect(countries,ignore.case("island"))
```

Use `str_subset` to find all countries that begin with a capital `A`.
```{r}
str_subset(countries,"^A")
```

Find all countries that end with `stan`
```{r}
str_subset(countries,"stan$")
```

Use `str_replace_all` to replace the `$`'s in `baws` with `s`'s.  Recall that the `$` is a metacharacter.
```{r}
baws <- "Bo$$"
str_replace_all(baws, "\\$", "s")
```

Use a character class (denoted by square brackets) to find all countries beginning with a vowel and ending with a consonant.
```{r}
begin_vowel<-"^[AEIOU].*[^aeiouAEIOU]$"
str_subset(countries,begin_vowel)
```




