---
title: "lab3"
author: "Kevin Khuu"
date: "June 8, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing the Data
Run install the package `readr` by running `install.packages("readr")` in the console.

The code below reads in the `tents.csv` comma-separated file.  The `col_types` argument is telling `R` what the data type is in each column.  See `?read_csv` if you're curious.
```{r}
library(readr)
tents <- read_csv("tents.csv", col_types = "ccnniccc")
```

### Questions
Answer the following questions programatically.  Some questions ask for you to explain what you see---you may answer this as text outside a code block.

1. Run `str()` on `tents`.  What information does it tell you? This is a useful function to know.
```{r}
str(tents)
```
I think that the str() gives you the structure of the input.

2. What class is `tents`?
```{r}
class(tents)
```
Class is either a dataframe,tbl or tbl_df.

3. How many rows are in `tents`?
```{r}
nrow(tents)
```
There are 90 rows in 'tents'

4. How many columns are in `tents`?
```{r}
ncol(tents)
```
There are 8 columns or variables in 'tents'

5. What are the names of the columns?
```{r}
names(tents)
```
6. Display the first 5 rows.

```{r}
tents[1:5,]
```

7. Display the last 4 rows.

```{r}
tents[87:90,]
```

8. Find out how to use `quantile` to find the interquartile range (75th percentile - 25th percentile) of the weights of the tents.

```{r}
quantile(tents$weight,prob=.75)-quantile(tents$weight,prob=.25)
```

9. Find the standard deviation of the heights of the `tents`. Does `R` use `n` or `n-1` in the denominator when calculating standard deviation?

```{r}
sd(tents$height)
help(sd)
```

R uses denominator n-1

10. The weight is given in grams.  Add a new column to the data frame called `weight_lbs` that contains the weight of the tents in pounds.  Remember to overwrite `tents` with your update.

```{r}
tents$weight_lbs<- round(tents$weight*0.00220462,2)
```

11. The height is given in centimeters.  Add a new variable `height_in` to the data frame that contains the height of the tents in inches. Remember to overwrite `tents` with your update.

```{r}
tents$height_in<- round(tents$height*0.393701,2)
```

12. How many tents cost more than $400?

```{r}
tents_more_400<- subset(tents,tents$price>400)
length(tents_more_400$price)
```

13. Subset the data frame so that it only contains tents that are valued $300 or less.  Assign the subsetted data frame the name `so_cheap_tents`.

```{r}
so_cheap_tents<- subset(tents,tents$price<300)
```

14. Of the tents that are more than $300, what percent are `big-agnes` tents?

```{r}
tents_more_300<- subset(tents, tents$price>300)
(sum(tents_more_300$brand=='big-agnes')/length(tents_more_300$brand))*100
```

15. Select the `bestuse` column of `tents` in four different ways: using `$`, `[[]]`, position number and `[]`, and variable name and `[]`.

```{r}
tents$bestuse
tents[["bestuse"]]
tents[6]
tents["bestuse"]
```

16. What is the name of the tend with the best height-to-weight ratio?

```{r}
tents$name[which.max(tents$height/tents$weight)]
```

17. Run `table` with two arguments, `seasons` and `capacity` (this should be one table).  What did it do?

```{r}
table(tents$season,tents$capacity)
```

It returns a contingency table of the seasons and capacity columns.

18. What type of tent (bestuse) is the most common? Your answer should return only the use and not the count.

```{r}
names(sort(table(tents$bestuse),decreasing=TRUE))[1]
```

19. Create a data frame called `kelty` with only data on kelty tents.  Do the same for `big-agnes`.  Then, combine the two data frames.

```{r}
kelty<- subset(tents,tents$brand=='kelty')
big_agnes<- subset(tents,tents$brand=='big-agnes')
combined_tents<-rbind(kelty,big_agnes)
combined_tents
```



